---
title: 'Homework 2 - Research in Health Economics'
author: "Ammarah Ahmed"
date: ""
header-includes:
  - \usepackage{booktabs}
output: 
  bookdown::pdf_document2:
    fig_caption: yes
    toc: false
---

```{r=setup, include=FALSE}
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, stringr, readxl, data.table, gdata, readr,scales, kableExtra, lemon, imputeTS, ivreg)
```

```{r, warning = FALSE, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE, fig.width = 16/2, fig.height = 9/2, tidy.opts=list(width.cutoff=60), tidy=TRUE)
```

```{r, include=FALSE}
final.data <- readRDS("data/output/TaxBurden_Data.rds")
```

# Summarise the Data

## Question 1

```{r, include=FALSE}
states <- unique(final.data$state) 

states_list <- list()
for (i in states) {
  state_data <- filter(final.data, state == i)
  state_data2 <- state_data %>% mutate(tax_change = c(0,ifelse(diff(tax_state) != 0, 1, 0)))
  states_list[[i]] <- state_data2
}
```

```{r, include=FALSE}
final.data2<- do.call(rbind, states_list)
```

```{r, include=FALSE}
tab_1 <- final.data2%>% filter(Year <= 1985) %>% 
  group_by(Year)%>% 
  summarize(prop = sum(tax_change == 1)/n())
```

```{r, include=FALSE}
fig_1<- ggplot(tab_1, aes(Year, prop)) + 
  geom_bar(stat = 'identity', colour="black") + 
  labs(title = "Proportion of States with Change in Tax, 1970-1985", x = "Year", y = "Proportion of States") +
  theme_bw()
```

```{r}
fig_1
```

## Question 2

```{r, include=FALSE}
tab_2 <- final.data %>% filter(Year <= 2018) %>% 
  group_by(Year)%>%
  summarise(avg_price = mean(cost_per_pack), avg_tax = mean(tax_dollar))
```
```{r, include=FALSE}
fig_2 <- ggplot(tab_2, aes(x = Year)) + 
  geom_line(aes(Year, avg_price), colour = "red") + 
  geom_line(aes(Year, avg_tax), colour = 'blue') +
  labs(title = "Average Price and Tax from 1970-2018", x = "Year", y = "Average Price and Tax") +
  theme_bw()
```

```{r}
fig_2
```

## Question 3

```{r, include=FALSE}
cig_data_diff <- final.data %>%
  group_by(state) %>%
  summarise(price_diff = cost_per_pack[Year == mean(2018)] - cost_per_pack[Year == mean(1970)]) %>%
  arrange(desc(price_diff))
```
```{r, include=FALSE}
top_5 <- cig_data_diff %>%
  slice(1:5)
```
```{r, include=FALSE}
tab_3<- final.data %>% filter (state %in% top_5$state & Year %in% c(1970:2018))%>% 
  group_by(Year, state)%>% 
  summarize(avg_sales = mean(sales_per_capita))
```
```{r, include=FALSE}
fig_3 <- ggplot(tab_3, aes(Year, avg_sales))+
  geom_line(aes(color = state))+
  labs(title = "Average Sales per Capita", x = "Year",y = "Average Sales per Capita")+
  theme_bw()
```

```{r}
fig_3
```

## Question 4

```{r, include=FALSE}
bottom_5 <- cig_data_diff %>%
  slice_min(price_diff, n = 5, with_ties = FALSE)
```
```{r, include=FALSE}
tab_4<- final.data %>% filter (state %in% bottom_5$state & Year %in% c(1970:2018))%>% 
  group_by(Year, state)%>% 
  summarize(avg_sales = mean(sales_per_capita))
```
```{r, include=FALSE}
fig_4 <- ggplot(tab_4, aes(Year, avg_sales))+
  geom_line(aes(color = state))+
  labs(title = "Average Sales per Capita", x = "Year",y = "Average Sales per Capita")+
  theme_bw()
```

```{r}
fig_4
```

## Question 5

The average cigarette sales fell more drastically between 1970 and 2018 for the states with the highest change in tax compared to states with the lowest change in tax. There was a more gradual decline in average sales for the latter states, with sales decreasing from a peak of ~140 packs to ~40 by 2018, except for North Carolina which had a sharper decline. In comparison, for states with highest change in tax, average sales fell from ~125 to ~10 packs in 2018.

# Estimate ATEs

## Question 1 
```{r, include=FALSE}
data_6 <- final.data %>% filter(Year %in% c(1970:1990))
```
```{r, include=FALSE}
data_6$log_sales <- log(data_6$sales_per_capita)
data_6$log_price <- log(data_6$cost_per_pack)

```
```{r, include=FALSE}
reg6<- lm(formula = log_price ~ log_sales, data = data_6)
```
```{r}
reg6
```

The regression shows that cigarette demand has an elasticity of -0.7335, which is inelestic, suggesting that cigarette demand is less responsive to changes in price.

## Question 2
```{r, include=FALSE}
data_7<- data_6 %>% mutate(total_tax= data_6$tax_dollar + data_6$tax_state)
  
```
```{r, include=FALSE}
iv7 <- ivreg(log_price ~ log_sales | total_tax, data = data_7)
```
```{r}
iv7
```
The IV regression looks wrong so I am unable to intepret results properly. 

## Question 3
need to fix previous errors to answer
## Question 4
need to fix previous errors to answer
## Question 5
need to fix previous errors to answer